---
title: "p8105_hw2_sc4456"
author: "Siyan Chen"
date: "9/29/2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
```
# Problem 1
```{r}
NYC_transit_data_modified = read_csv("./hw2_data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada) %>% 
  mutate(entry = recode(entry, `YES`= "TRUE", `NO` = "FALSE" )) 
NYC_transit_data_modified
```
* The new dataset contains viables of line, station name, station latitude, station longitue, 1 to 11 routs served, entry, vending, entrance type and ADA notes. For the data cleaning steps, I import the original dataset and modify the variable names and then use select function to retain the variables I interested. The new dataset gives the `r nrow (NYC_transit_data_modified)` rows and `r ncol(NYC_transit_data_modified)`. This is not tidy data.
```{r}
distinct(NYC_transit_data_modified,station_name,line) %>% 
  nrow()
# Firstly, I use distinct function to find the distinct station which excludes the repeated stations. Then, I use nrow function to fund the number of distinct stations 
NYC_transit_data_modified %>% 
  filter(ada == "TRUE") %>% 
  distinct(station_name,line) %>% 
  nrow()
# This is to get the anwser of how many stations are ADA complaint. Firstly, I removed the NA data of ADA notes.Then I use distinct function to found our how many distinct stations exist because some ADA compliants target the same station. It returns the 84 stations are ADA compliant. 
```

* There are `r nrow(distinct (NYC_transit_data_modified, line, station_name))` distinct stations

* There are 84 stations are ADA compliant.

```{r}
NYC_transit_data_modified %>% 
filter(vending == "YES") %>% 
  nrow()/nrow(NYC_transit_data_modified)
```
* What proportion of station entrances / exits without vending allow entrance? It returns 90.2% station entrances/exits without vending allow entrance. 

```{r}
NYC_transit_data_modified %>% 
  gather(., key = "route_number", value = "route_name", 5:15) %>% 
  filter(., route_name == "A") %>% 
  distinct(., station_name, line) %>% 
  nrow() 
NYC_transit_data_modified
# I use gather function to reformat date from wide to long. To get the number of distinct stations serving A train, I filter the data to only keep the rows with A trains. Then I use previous method to get the number of distinct stations, which returns 60. 
```

```{r}
NYC_transit_data_modified %>% 
  gather(., key = "route_number", value = "route_name", 5:15) %>% 
  filter(., route_name == "A") %>% 
  filter(ada == "TRUE") %>% 
  distinct(., station_name, line) %>% 
  nrow() 

# Use filter function to firstly keep the row with A train and use filter function again to keep the row with ada complaints. It returns 17.
```
* There are 56 distinct stations serve the A train.Of the stations that serve the A train, there are 17 have ADA complaint.

# Problem 2

```{r}
trashwheel_data = readxl::read_xlsx("./hw2_data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "Mr. Trash Wheel") %>% 
  janitor::clean_names() %>% 
  select(-x_1) %>% 
  filter(!is.na(dumpster)) %>% 
  mutate(sports_balls = as.integer(round(sports_balls)))
trashwheel_data
```
```{r}
precipitation2016 = readxl::read_xlsx("./hw2_data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "2016 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  filter(!is.na(total)) %>% 
  mutate(year = "2016") %>% 
  select(year, everything())
  
precipitation2017 = readxl::read_xlsx("./hw2_data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "2017 Precipitation", range = "A2:B14")%>% 
  janitor::clean_names() %>% 
  filter(!is.na(total)) %>% 
  mutate(year = "2017") %>% 
  select(year, everything()) 

precipitation_tidy = bind_rows(precipitation2016, precipitation2017) %>%   
  mutate(month = month.name[month]) 
precipitation_tidy

```
There are `r nrow(precipitation2016)` observations in 2016 and `r nrow(precipitation2017)` observation in 2017. The Key Variable includs year, month and total precipitation. The total precipitation in 2017 is `r sum(precipitation2017$total)` in. The median number of sports balls in a dumpster is `r median(trashwheel_data$sports_balls)`

# Problem 3
```{r}
devtools::install_github("p8105/p8105.datasets")
library(p8105.datasets)
data(brfss_smart2010)
modified_brfss =
brfss_smart2010 %>% 
janitor::clean_names() %>% 
filter(topic == "Overall Health") %>% 
select(-class, -topic, -question, -sample_size, -confidence_limit_low : -geo_location) %>% 
spread(response, data_value) %>% 
janitor::clean_names() %>% 
mutate(excellent_or_very_good = excellent + very_good) 
```
```{r}
modified_brfss %>% 
  distinct(locationdesc) %>% 
  nrow()
# It returns the 404 unique locations
modified_brfss %>% 
  distinct(locationabbr) %>% 
  nrow()
# It returns 51 states, so every state is representated
names(which.max(table(modified_brfss$locationabbr)))
# This method is what I found on google. I am trying to understand how it works. I guess it first help to fing the maximum occuring character and gives the name of the character. I guess the reason to use table function is that which.max function only works for table. It returns NJ.
```

* There are  and there are 51 distinct states, so every state is represented. `r names(which.max(table(modified_brfss$locationabbr)))` state is observed the most
* The median of Excellent value is `r median(modified_brfss$excellent, na.rm = TRUE)`
* Histogram 
```{r}
modified_brfss_2002 = filter(modified_brfss, year == "2002")
# get the data of year of 2002
ggplot(modified_brfss_2002, aes(x = excellent)) + geom_histogram()
```
* Scatterplot
```{r}
newyork_county_queens_county_data = filter(modified_brfss, locationdesc == "NY - Queens County" | locationdesc =="NY - New York County")
# get the data of New York county and Queens county
ggplot(newyork_county_queens_county_data, aes(x = year, y = excellent))+geom_point(aes(color = locationdesc))

```







